Ignition of Fuel Beds by Firebrands

### 2022-04-05 litreview.tex
Continued development of the background/literature review

    \section{Historical Perspective}
    Wild fires have been an integral part of many ecosystems for millennia and are often beneficial or required for maintaining a healthy ecosystem. As beneficial as they may be to the natural enviornment wildfires pose significant risk to the built environment. As communities around the world have expanded communities outward into environments that support wildfires (e.g., forests and grasslands) the risk of destruction of structures by wildfires has increased. In the United States 91\% of the Wildland Urban Interface \nomenclature{WUI}{Wildland Urban Interface} growth of 189,000\si{\kilo\meter\squared} (larger than the state of Washington) development between 1990 and 2010 is occupied by homes~\cite{Radeloff2017} and has continued to grow on recent years. 
    \section{The ignition process}

>>>>

    Wild fires have been an integral part of many ecosystems for millennia and are often beneficial or required for maintaining a healthy ecosystem. As beneficial as they may be to the natural enviornment wildfires pose significant risk to the built environment. As communities around the world have expanded communities outward into environments that support wildfires (e.g., forests and grasslands) the risk of destruction of structures by wildfires has increased. In the United States 91\% of the Wildland Urban Interface \nomenclature{WUI}{Wildland Urban Interface} growth of 189,000\si{\kilo\meter\squared} (larger than the state of Washington) development between 1990 and 2010 is occupied by homes~\cite{Radeloff2017} and has continued to grow on recent years. The increased development in theses areas means that an excess of 40 million homes are located in areas that may be threatened by a wildfire.

    In addition to the increase of homes in the WUI the intensity and severity of fires has increased. California is a prime example of the increased fire severity where 13 of the 20 most destructive fires have occurred since 2010 and 7 of the 20 most destructive occurred in 2020 and 2021~\cite{CALFIRE2018}. Increasingly severe fires are not just limited to California or the United States. Texas and Tennessee each had one of the top ten most severe fires in the United States between 2005 and 2020 and Colorado, Oregon, Washington, Oklahoma, and Florida have also seen highly destructive fires in recent years~\cite{Barrett2020}. The United States is not unique in the presence of wilfdires that destroy homes. Australia's Black Summer in 2019-2020 was one of the most severe in the countries history~\cite{Levin2021}. The Mediterranean countries of Portugal, Spain, France, Italy, and Greece are also either experiencing greater losses or are a risk for greater losses due to wildfires in recent years. Similar trends are also present in many Asian and South American countries~\cite{Manzello2018}. Wildfires that threaten lives, homes, and structures are truly a global phenomena and it is essential to understand how homes are ignited and destroyed in wildfires to better prevent losses from occurring.

    Ignition of a home in a WUI fire is attributed to three primary mechanisms; direct flame contact, ignition by radiative heating, and ignition by firebrands. Of these methods ignition by firebrands is considered to be a primary threat to homes~\cite{Suzuki2021, Mell2010}. A case study of structure ignitions in a WUI fire found that firebrand ignition was responsible for 2/3 of structure ignitions~\cite{Mell2011}. With the majority of ignitions occurring due to firebrand ignition it is imperative to understand the mechanism of firebrand ignition to reduce home losses. The general mechanism for firebrand ignition is considered to be a three step process which starts within the fire itself, often at a significant distance from the ignition point. First, a firebrand is generated as burning material breaks off or is lofted into the air. The firebrand is then lofted through the air and transported to the ignition site. Upon landing on a recipient fuel bed ignition occurs if the criteria for ignition is met~\cite{Babrauskas2003}. When considering the possibility of ignition during a fire event all three of processes must be considered as each process significantly influences the properties of an ember landing on a fuel bed. Due to the highly variable nature of winds that transport firebrands during fires the firebrand the firebrand generation and fuel bed ignition processes may be the most consistent when considering prevention approaches. Foliage types are much more consistent and measureable than wind and have a significant impact on the type of embers produced~\cite{Hudson2020EffectsScale} and fuel beds are largely controlled and/or modifiable by homeowners. Unfortunately, the generation and ignition processes are significantly less understood than the transport processes~\cite{Manzello2020}. The work contained in this dissertation is focused on the ignition process of the fuel bed by a firebrand and does not consider the generation of firebrands. For further information on firebrand generation processes the reader is referred to review papers considering this process by Fernandez-Pello~\cite{Fernandez-Pello2017} and by Manzello et al.~\cite{Manzello2020}.

    Developing and increased understanding of the ignition of a fuel bed has been identified as a necessary step to better protect structures from ignition for decades but 
    \section{The ignition process}
